### 2020-07-16 article.tex
Remarks of Jose

\section{KSPHPDDM}\label{sec:KSPHPDDM}
  \subsection{Related work}
Krylov subspace methods are widely used in numerical linear algebra for solving
linear systems of equations because of their memory efficiency~\cite{barrett1994templates}.
They mainly rely on matrix--vector
operations such as multiplications or transpose multiplications and,
for robust convergence when dealing with challenging high-dimensional systems,
on the application of appropriate preconditioners.
PETSc (as of version 3.13.2) already offers 44 different types of Krylov methods within the \pk{KSP} base class. Some of the
most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
the generalized minimal residual method~\cite{saad1986gmres}~(GMRES), conjugate
gradient~\cite{hestenes1952methods}~(CG), or biconjugate gradient stabilized method~\cite{van1992bi}~(BCGS). % Some of these methods are also
% available in other linear algebra backends such as~\emph{hypre} or DUNE~\cite{bastian2010generic}.

@@

According to the standard PETSc notation, ksp($A$, $P_{A'}$)
defines a \pk{KSP} used to solve a system $Ax=b$ using a preconditioner $P_{A'}$ built
using an operator $A'$ which is usually, but not necessarily, equal to $A$. While solving
the linear system via \pk{KSPSolve}(ksp, $b$, $x$) or \pk{KSPMatSolve}(ksp, $B$, $X$),
HPDDM will repeatedly call the following PETSc routines:
\begin{itemize}
    \item \pk{MatMult}($A$, $x$, $y$) for $y = Ax$;
    \item \pk{MatMatMult}($A$, $X$, {\small \pk{MAT\_REUSE\_MATRIX}}, {\small
        \pk{PETSC\_DEFAULT}}, $Y$)\footnote{or \pk{MatProductNumeric}($Y$) with PETSc 3.14.0 and above} for $Y =
        AX$;
    \item \pk{PCApply}($P_{A'}$, $y$, $x$) for $x = P_{A'}y$;
    \item \pk{PCMatApply}($P_{A'}$, $Y$, $X$) for $X = P_{A'}Y$;
\end{itemize}
As of version 3.14.0, PETSc provides unified routines for applying $A$ or $P_{A'}$
to a tall-and-skinny dense matrix $Y$. However, when no specialized implementation
is available, PETSc will loop over the columns of $Y$ and successively apply $A$
or $P_{A'}$ to get all columns of $X$. For matrix--matrix products, there are
currently specialized implementations for the following \pk{MatType}:
\begin{itemize}
    \item \pk{MatMPIAIJ};
    \item \pk{MatShell};
    \item \pk{MatNest};
    \item \pk{MatSeqAIJ}, \pk{MatSeqBAIJ}, and \pk{MatSeqSBAIJ};
    \item \pk{MatSeqAIJcuSPARSE}.
\end{itemize}
The following preconditioners have specialized implementations for dealing
efficiently with multiple vectors:
\begin{enumerate}
    \item \pk{PCNONE}: $P_{A'} = I$, identity matrix;
    \item \pk{PCKSP}: embedded Krylov method inside $P_{A'}$;
    \item \pk{PCMAT}: multiplication by $A'$, i.e., $P_{A'} = {A'}$;
    \item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
    \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
    \item \pk{PCHARA}~\cite{10.1145/3232850};
    \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
    \item \pk{PCASM} and \pk{PCGASM}: overlapping Schwarz methods;\label{it:asm}
    \item \pk{PCBJACOBI}: block Jacobi;\label{it:bjacobi}
    \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization,
        i.e., $P_{A'} = {A'}^{-1}$;\label{it:lu}
    \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky
        factorization.\label{it:ilu}
\end{enumerate}
One appealing feature of domain decomposition-like preconditioners 
\cref{it:hpddm,it:asm,it:bjacobi} is that they most often rely on exact or inexact
factorizations~\cref{it:lu,it:ilu} as subdomain solvers.
In these cases, it is possible to access the so-called local or distributed factored matrix $F$
%via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $Y$, $X$) to take advantage
of blocked forward eliminations and backward substitutions from the various factorization packages
interfaced with PETSc. In the case of a sparse coefficient matrix $A'$,
this strategy is possible with: MUMPS~\cite{amestoy2001fully},
SuiteSparse~\cite{davis2004algorithm}, MKL
PARDISO or CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense coefficient
matrix $A'$: ScaLAPACK and Elemental~\cite{poulson2013elemental}. 
Future work will consider extending the various implementations of PETSc
preconditioning infrastructure to increase the arithmetic intensity of the
block preconditioner application phase.

The rest of the operations are then performed directly inside HPDDM.
Specifically, within Krylov methods using the Arnoldi process, or when recycling is requested,
one has to compute $QR$ factorizations of tall-and-skinny dense matrices to orthonormalize
candidate basis vectors. Such operations are performed using the CholQR algorithm~\cite{stathopoulos2002block}, or
via the block variants of the (modified) Gram--Schmidt method.
For Hessenberg matrices generated by the Arnoldi
process, it is common to update their $QR$ decomposition using Givens
rotations~\cite{saad2003iterative}. For block Hessenberg matrices, Householder reflectors are
used instead~\cite{gutknecht2008updating}.
These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. {\color{red} API?}

    \subsubsection{User-defined deflation space} Recycling Krylov methods try to
extract convergence information by solving a small standard or a generalized
dense eigenproblem at the end of each cycle or when convergence is reached, and
then reuse it appropriately for subsequent
solves~\cite{saad2000deflated,stathopoulos2002block,soodhalter2014krylov}.
The routine \pk{KSPHPDDMGetDeflationSpace} can be used to retrieve such deflation subspaces, stored as distributed
dense tall-and-skinny matrices. It is also possible to provide this information
via \pk{KSPHPDDMSetDeflationSpace} in case end-users already know a good initial
deflation subspace.

@@

As a side note, when dealing with
\pk{MatKAIJ} matrices, for which it is assumed that an operator may be written as $I \otimes S
+ A \otimes T$, with $S$ and $T$ elements of $\mathbb{K}^{p\times q}$ and
$A$ a standard sparse matrix of dimension $n$, it is also possible to use block Krylov methods. 
Indeed, if $p=q=k$, $S = \lambda I$, and $T =
\mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
system with multiple right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
the row-major representation of $b$, i.e., $b_{k \cdot i + j} = B_{i, j},
\forall (i, j) \in \llbracket 1;n\rrbracket \times \llbracket 1;k\rrbracket$.
On the one hand, due to the structure of the operator, it is challenging to
define appropriate preconditioners for $(I \otimes S + A \otimes T)$: only
\pk{PCPBJACOBI} is currently implemented for \pk{MatKAIJ}. On the other hand,
it may be simpler to efficiently approximate the inverse of $A$. In
\pk{KSPHPDDM}, the type of the operators is checked: if it is \pk{MatKAIJ} and
the previous condition on $p, q, S$, and $T$ is fulfilled, the single
right-hand side system is automatically cast into a system with multiple
right-hand sides instead.

@@

processes and they can be launched using the following command line: \\[-4pt]

@@

&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -sub_pc_type icc -ksp_rtol 1e-10          &fvtextcolor[myred][\]

@@

@@

The iteration numbers needed to reach convergence are gathered in~\cref{fig:comparison};
the cases \ref{pgfplots:r0} and

@@

401 498         293        457           497        131       460          127       438        127       438
402 228         125        187           231        53        184          127       449        127       449
403 209         113        182           206        47        177          127       450        127       450
404 199         112        175           198        47        174          127       450        127       450
405 198         110        174           198        47        172          127       451        127       451
406 198         111        174           199        47        174          127       452        127       452
407 206         114        180           206        49        179          126       458        126       458
408 207         115        181           208        49        182          128       458        135       458
409 207         110        181           207        49        181          128       458        128       458
410 206         110        179           206        49        179          128       457        128       457

@@

evaluated using a given base flow $q_b = \begin{bmatrix}u_b & p_b\end{bmatrix}^T$ and

@@

The effectiveness of subspace recycling is shown for
finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands.
We note here that the first two commands are merely used to generate the base flow $q_b$ with \pk{SNESSolve} using a
continuation method on the Reynolds number. \\[-4pt]
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 50 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 100 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type fgmres -st_ksp_gmres_restart 200 -st_ksp_converged_reason
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type hpddm -st_ksp_gmres_restart 200 -st_ksp_converged_reason              &fvtextcolor[myred][\]
  -st_ksp_hpddm_variant flexible -st_ksp_hpddm_recycle 10 -st_ksp_hpddm_type gcrodr
\end{Verbatim}

@@

The readiness of PETSc, as of version 3.13.2, is now discussed

@@

In particular, we consider three PETSc matrix types:
\begin{itemize}
    \item \pk{MatSeqAIJ}, standard sequential sparse matrices, based on compressed sparse row format;
    \item \pk{MatSeqBAIJ}, block sparse matrices, based on block compressed sparse row format;
    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrices stored in upper triangular form.
\end{itemize}
Their distributed memory counterparts rely on the sequential implementations, with each process
storing two such matrices, one with local rows and columns, and another with local rows and ``nonlocal'' columns.
To keep this study succinct, we only evaluate the performance of the sequential implementations, which can help
in drawing conclusions for the intraprocess performance of the parallel formats.
For this reason, from now on, the \pk{Seq} substring is dropped.
The performance of the multiplication primitives of the three aforementioned types will also be compared to
those obtained with the MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
There is an ongoing effort to better integrate these routines in PETSc, but at the time of writing, direct calls to the MKL were used.

@@

discretizing partial differential equations. We then generate a random symmetric dense matrix $T$ of dimension $b$,

@@

They can be reproduced by using the mini-app \url{https://github.com/prj-/jolivet2020petsc/blob/master/MatProduct.c} and any input \pk{MatAIJ} stored in binary format,

@@

The mini-app is then used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and using cuSPARSE~\cite{cusparse-web-page} as interfaced in PETSc.
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][export] OMP_NUM_THREADS=20 &&&& &fvtextcolor[myyellow][export] MKL_NUM_THREADS=20
  -bs 1,3,6 -N 1,2,8,16,32,64 -type aijmkl,baijmkl,sbaijmkl,aijcusparse
Results reported in~\cref{fig:perfAXpar} have been scaled with respect to the
baseline implementation \pk{MatAIJMKL} with $N=1$. They have been obtained on a
single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math
mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at
\SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2. Neither transfers
between host and device nor time spent in \texttt{mkl\_sparse\_optimize} have
been taken into account. For the sake of completeness, the performance of
\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported in
\cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
perform this type of workload for such small sparse matrices and skinny dense
matrices. But this helps in drawing a fair comparison between a full CPU socket
and a full GPU device. All in all, the CPU configuration which reaches the
highest percentage of peak is MatSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
For the GPU configuration, \pk{MatAIJcuSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2.
P       aij1        aij3        aij6
1       6.6857e-02  7.4649e-01  3.1128e+00       
2       3.4206e-02  3.7820e-01  1.5723e+00
4       1.6565e-02  1.8634e-01  7.7238e-01
8       9.1370e-03  9.7549e-02  3.9884e-01
16      7.2334e-03  7.9871e-02  3.2154e-01
20      7.8399e-03  8.6487e-02  3.5536e-01
}\loadedOMP
\pgfplotstablegetelem{0}{aij1}\of{\loadedOMP}
\let\refI\pgfplotsretval
\pgfplotstablecreatecol[create col/expr={\refI/\thisrow{aij1}}]{speedupI}\loadedOMP
\pgfplotstablegetelem{0}{aij3}\of{\loadedOMP}
\let\refIII\pgfplotsretval
\pgfplotstablecreatecol[create col/expr={\refIII/\thisrow{aij3}}]{speedupIII}\loadedOMP
\pgfplotstablegetelem{0}{aij6}\of{\loadedOMP}
\let\refVI\pgfplotsretval
\pgfplotstablecreatecol[create col/expr={\refVI/\thisrow{aij6}}]{speedupVI}\loadedOMP
      \begin{loglogaxis}[
              ytick={1,2,4,8,20},
            \pgfmathparse{100*round(exp(\tick))}
        extra x ticks = {1,2,4,8,16,20},height=6cm,
          extra x tick labels = {1,2,4,8,16,\hspace*{0.3cm}20},
          legend style={at={(0.02,0.98)}, legend cell align=left,anchor=north west},
          \addplot[black, dashed] table [x expr=\thisrow{P},y expr=\thisrow{P}]{\loadedOMP};\label{pgfplots:linear}
          \addplot[blueaccent,mark=square*] table [x expr=\thisrow{P},y expr=\thisrow{speedupI}]{\loadedOMP};\label{pgfplots:speedupI}
          \addplot[darkgray,mark=*] table [x expr=\thisrow{P},y expr=\thisrow{speedupIII}]{\loadedOMP};\label{pgfplots:speedupIII}
          \addplot[last,mark=triangle*] table [x expr=\thisrow{P},y expr=\thisrow{speedupVI}]{\loadedOMP};\label{pgfplots:speedupVI}
          \addlegendentry{Linear speedup}
          \addlegendentry{$b = 1$}
          \addlegendentry{$b = 3$}
          \addlegendentry{$b = 6$}
    \end{loglogaxis}
    \caption{Scalability of \pk{MatMult} with \pk{MatAIJMKL} and different number of processes.\label{fig:scaling-Ax}}

@@

           \addlegendentry{\pk{MatAIJcuSPARSE}}

@@

@@

decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied local unassembled
matrices. These matrices are equivalent to the $\{\mathring{A}_i\}_{i=1}^N$

@@

\item supply whenever the Jacobian is being updated in the \pk{SNESSetJacobian}
function new unassembled Neumann matrices at the current linearization point with the previously
explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically with a

@@

\bibitem{grote1997parallel}
{\sc M.~J. Grote and T.~Huckle}, {\em Parallel preconditioning with sparse
  approximate inverses}, SIAM Journal on Scientific Computing, 18 (1997),
  pp.~838--853.

>>>>

\section{KSPHPDDM}\label{sec:KSPHPDDM}
  \subsection{Related work}
Krylov subspace methods are widely used in numerical linear algebra for solving
linear systems of equations because of their memory efficiency~\cite{barrett1994templates}.
They mainly rely on matrix--vector
operations such as multiplications or transpose multiplications and,
for robust convergence when dealing with challenging high-dimensional systems,
on the application of appropriate preconditioners.
PETSc, as of version 3.14.0, already offers 44 different types of
Krylov methods within the \pk{KSP} base class. Some of the
most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
the generalized minimal residual method~\cite{saad1986gmres}~(GMRES), conjugate
gradient~\cite{hestenes1952methods}~(CG), or biconjugate gradient stabilized method~\cite{van1992bi}~(BCGS). % Some of these methods are also
% available in other linear algebra backends such as~\emph{hypre} or DUNE~\cite{bastian2010generic}.

@@

According to the standard PETSc notation, ksp($A$, $P_{A'}$)
defines a \pk{KSP} used to solve a system $Ax=b$ using a preconditioner $P_{A'}$ built
using an operator $A'$ which is usually, but not necessarily, equal to $A$. While solving
the linear system via \pk{KSPSolve}(ksp, $b$, $x$) or \pk{KSPMatSolve}(ksp, $B$, $X$),
HPDDM will repeatedly call the following PETSc routines:
\begin{itemize}
    \item \pk{MatMult}($A$, $x$, $y$) for $y = Ax$;
    \item \pk{MatMatMult}($A$, $X$, {\small \pk{MAT\_REUSE\_MATRIX}}, {\small
        \pk{PETSC\_DEFAULT}}, $Y$)\footnote{or \pk{MatProductNumeric}($Y$) with PETSc 3.14.0 and above} for $Y =
        AX$;
    \item \pk{PCApply}($P_{A'}$, $x$, $y$) for $y = P_{A'}x$;
    \item \pk{PCMatApply}($P_{A'}$, $X$, $Y$) for $Y = P_{A'}X$;
\end{itemize}
As of version 3.14.0, PETSc provides unified routines for applying $A$ or $P_{A'}$
to a tall-and-skinny dense matrix $X$. However, when no specialized implementation
is available, PETSc will loop over the columns of $X$ and successively apply $A$
or $P_{A'}$ to get all columns of $Y$. For matrix--matrix products, there are
currently specialized implementations for the following \pk{MatType}:
\begin{itemize}
    \item \pk{MatSeqAIJ}, standard sequential sparse matrix, based on compressed sparse row format;
    \item \pk{MatSeqBAIJ}, block sparse matrix, based on block compressed sparse row format;
    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrix stored in upper triangular form.
    \item \pk{MatMPIAIJ}, standard parallel sparse matrix;
    \item \pk{MatShell}, user-defined matrix;
    \item \pk{MatNest}, block-defined matrix with nested submatrices;
    \item \pk{MatSeqAIJCUSPARSE}, standard sequential sparse matrix, offloaded to a GPU using cuSPARSE~\cite{cusparse-web-page}.
\end{itemize}
The following preconditioners have specialized implementations for dealing
efficiently with multiple vectors:
\begin{enumerate}
    \item \pk{PCNONE}: $P_{A'} = I$, identity matrix;
    \item \pk{PCKSP}: embedded Krylov method inside $P_{A'}$;
    \item \pk{PCMAT}: multiplication by $A'$, i.e., $P_{A'} = {A'}$;
    \item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
    % \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
    \item \pk{PCHARA}~\cite{10.1145/3232850};
    \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
    \item \pk{PCASM} and \pk{PCGASM}: overlapping Schwarz methods;\label{it:asm}
    \item \pk{PCBJACOBI}: block Jacobi;\label{it:bjacobi}
    \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization,
        e.g., $P_{A'} = {L}^{-1}{U}^{-1}$;\label{it:lu}
    \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky
        factorization.\label{it:ilu}
\end{enumerate}
One appealing feature of domain decomposition-like preconditioners 
\cref{it:hpddm,it:asm,it:bjacobi} is that they most often rely on exact or inexact
factorizations~\cref{it:lu,it:ilu} as subdomain solvers.
In these cases, it is possible to access the so-called local or distributed factored matrix $F$
%via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $X$, $Y$) to take advantage
of blocked forward eliminations and backward substitutions from the various factorization packages
interfaced with PETSc. In the case of a sparse coefficient matrix $A'$,
this strategy is possible with: MUMPS~\cite{amestoy2001fully},
SuiteSparse~\cite{davis2004algorithm}, MKL
PARDISO or CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense coefficient
matrix $A'$: ScaLAPACK and Elemental~\cite{poulson2013elemental}. 
Future work will consider extending the various implementations of PETSc
preconditioning infrastructure to increase the arithmetic intensity of the
block preconditioner application phase.

The rest of the operations are then performed directly inside HPDDM.
Specifically, within Krylov methods using the Arnoldi process, or when recycling is requested,
one has to compute $QR$ factorizations of tall-and-skinny dense matrices to orthonormalize
candidate basis vectors. Such operations are performed using the CholQR algorithm~\cite{stathopoulos2002block}, or
via the (modified) Gram--Schmidt method.
For Hessenberg matrices generated by the Arnoldi
process, it is common to update their $QR$ decomposition using Givens
rotations~\cite{saad2003iterative}. For block Hessenberg matrices, Householder reflectors are
used instead~\cite{gutknecht2008updating}.
These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. {\color{red} API?}

    \subsubsection{User-defined deflation space} Recycling Krylov methods try to
extract convergence information by solving a small standard or a generalized
dense eigenproblem at the end of each cycle or when convergence is reached, and
then reuse it appropriately for subsequent
solves~\cite{saad2000deflated,stathopoulos2009deflation,soodhalter2014krylov}.
The routine \pk{KSPHPDDMGetDeflationSpace} can be used to retrieve such deflation subspaces, stored as distributed
dense tall-and-skinny matrices. It is also possible to provide this information
via \pk{KSPHPDDMSetDeflationSpace} in case end-users already know a good initial
deflation subspace.

@@

As a side note, when dealing with
\pk{MatKAIJ} matrices, for which it is assumed that an operator may be written as $I \otimes S
+ A \otimes T$, with $S$ and $T$ elements of $\mathbb{K}^{p\times q}$ and
$A$ a standard sparse matrix of dimension $n$, it is also possible to use block Krylov methods. 
Indeed, if $p=q=k$, $S = \lambda I$, and $T =
\mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
system with $k$ right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
the row-major representation of $b$, i.e., $b_{k \cdot i + j} = B_{i, j},
\forall (i, j) \in \llbracket 1;n\rrbracket \times \llbracket 1;k\rrbracket$.
On the one hand, due to the structure of the operator, it is challenging to
define appropriate preconditioners for $(I \otimes S + A \otimes T)$: only
\pk{PCPBJACOBI} is currently implemented for \pk{MatKAIJ}. On the other hand,
it may be simpler to efficiently approximate the inverse of $A$. In
\pk{KSPHPDDM}, the type of the operators is checked: if it is \pk{MatKAIJ} and
the previous condition on $p, q, S$, and $T$ is fulfilled, the single
right-hand side system is automatically cast into a system with multiple
right-hand sides instead.

@@

processes and they can be launched using the following command lines: \\[-4pt]

@@

&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -redundant_pc_type icc -ksp_rtol 1e-10    &fvtextcolor[myred][\]

@@

  -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR -pc_type redundant

@@

For the case of ICC(0), a \pk{PCREDUNDANT} is needed as well since
PETSc does not provide a parallel ICC solver. The iteration numbers
needed to reach convergence are gathered in~\cref{fig:comparison}.  The
cases \ref{pgfplots:r0} and

@@

401 498         293        457           497        94        460          127       438        127       438
402 228         125        187           231        37        184          127       449        127       449
403 209         113        182           206        34        177          127       450        127       450
404 199         112        175           198        33        174          127       450        127       450
405 198         110        174           198        34        172          127       451        127       451
406 198         111        174           199        34        174          127       452        127       452
407 206         114        180           206        33        179          126       458        126       458
408 207         115        181           208        34        182          128       458        135       458
409 207         110        181           207        34        181          128       458        128       458
410 206         110        179           206        34        179          128       457        128       457

@@

evaluated using a given base flow $q_b = \begin{bmatrix}u_b \\ p_b\end{bmatrix}$ and

@@

The effectiveness of subspace recycling is shown for
finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands.
We note here that the first two commands are merely used to generate
the base flow $q_b$ with a \pk{SNES}, a PETSc object used to solve
nonlinear problems, using a continuation method on the Reynolds number. \\[-4pt]
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 50 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 100 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type fgmres -st_ksp_gmres_restart 200 -st_ksp_converged_reason
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type hpddm -st_ksp_gmres_restart 200 -st_ksp_converged_reason              &fvtextcolor[myred][\]
  -st_ksp_hpddm_variant flexible -st_ksp_hpddm_recycle 10 -st_ksp_hpddm_type gcrodr
\end{Verbatim}

@@

The readiness of PETSc, as of version 3.14.0, is now discussed

@@

In particular, the first three matrix types from~\cref{sec:interface-ksp}, \pk{MatSeqAIJ}, \pk{MatSeqBAIJ}, and \pk{MatSeqSBAIJ} are first considered.
Their distributed memory counterparts, e.g., \pk{MatMPIAIJ}, rely on the sequential implementations, with each process
storing two such matrices, one with local rows and columns, and another with local rows and ``nonlocal'' columns.
To keep this study succinct, we only evaluate the performance of the sequential implementations, which can help
in drawing conclusions for the intraprocess performance of the parallel formats.
For this reason, from now on, the \pk{Seq} substring is dropped.
The performance of the multiplication primitives of the three aforementioned types will also be compared to
those obtained with the MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
There is an ongoing effort to better integrate these routines in PETSc, but at the time of writing, direct calls to the MKL were used.

@@

discretizing partial differential equations. We then generate a random symmetric dense $b$-by-$b$ matrix $T$,

@@

They can be reproduced by using the mini-app \texttt{MatProduct.c} from \url{https://github.com/prj-/jolivet2020petsc} and any input \pk{MatAIJ} stored in binary format,

@@

The mini-app is then used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and using cuSPARSE~\cite{cusparse-web-page} as interfaced in PETSc.
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][export] OMP_NUM_THREADS=20 &&&& &fvtextcolor[myyellow][export] MKL_NUM_THREADS=20
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 1 ./MatProduct -f binaryoutput -log_view                                 &fvtextcolor[myred][\]
  -bs 1,3,6 -N 1,2,8,16,32,64 -type aijmkl,baijmkl,sbaijmkl,aijcusparse
\end{Verbatim}
Results reported in~\cref{fig:perfAXpar} have been scaled with respect to the
baseline implementation \pk{MatAIJMKL} with $N=1$. They have been obtained on a
single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math
mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at
\SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2. Neither transfers
between host and device nor time spent in \texttt{mkl\_sparse\_optimize} have
been taken into account.

@@

           \addlegendentry{\pk{MatAIJCUSPARSE}}

@@

For the sake of completeness, the performance of
\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported in
\cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
perform this type of workload for such small sparse matrices and skinny dense
matrices. But this helps in drawing a fair comparison between a full CPU socket
and a full GPU device. All in all, the CPU configuration which reaches the
highest percentage of peak is MatSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
For the GPU configuration, \pk{MatAIJCUSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2.
\pgfplotstableread{
P       aij1        aij3        aij6
1       6.6857e-02  7.4649e-01  3.1128e+00       
2       3.4206e-02  3.7820e-01  1.5723e+00
4       1.6565e-02  1.8634e-01  7.7238e-01
8       9.1370e-03  9.7549e-02  3.9884e-01
16      7.2334e-03  7.9871e-02  3.2154e-01
20      7.8399e-03  8.6487e-02  3.5536e-01
}\loadedOMP
\pgfplotstablegetelem{0}{aij1}\of{\loadedOMP}
\let\refI\pgfplotsretval
\pgfplotstablecreatecol[create col/expr={\refI/\thisrow{aij1}}]{speedupI}\loadedOMP
\pgfplotstablegetelem{0}{aij3}\of{\loadedOMP}
\let\refIII\pgfplotsretval
\pgfplotstablecreatecol[create col/expr={\refIII/\thisrow{aij3}}]{speedupIII}\loadedOMP
\pgfplotstablegetelem{0}{aij6}\of{\loadedOMP}
\let\refVI\pgfplotsretval
\pgfplotstablecreatecol[create col/expr={\refVI/\thisrow{aij6}}]{speedupVI}\loadedOMP
\begin{figure}[tbp]
  \centering
    \begin{tikzpicture}
      \begin{loglogaxis}[
              ytick={1,2,4,8,20},
        yticklabel={
            \pgfmathparse{100*round(exp(\tick))}
$\pgfmathprintnumber{\pgfmathresult}$\%
},
        xtick = {\empty},
        extra x ticks = {1,2,4,8,16,20},height=6cm,
          extra x tick labels = {1,2,4,8,16,\hspace*{0.3cm}20},
          legend style={at={(0.02,0.98)}, legend cell align=left,anchor=north west},
      ]
          \addplot[black, dashed] table [x expr=\thisrow{P},y expr=\thisrow{P}]{\loadedOMP};\label{pgfplots:linear}
          \addplot[blueaccent,mark=square*] table [x expr=\thisrow{P},y expr=\thisrow{speedupI}]{\loadedOMP};\label{pgfplots:speedupI}
          \addplot[darkgray,mark=*] table [x expr=\thisrow{P},y expr=\thisrow{speedupIII}]{\loadedOMP};\label{pgfplots:speedupIII}
          \addplot[last,mark=triangle*] table [x expr=\thisrow{P},y expr=\thisrow{speedupVI}]{\loadedOMP};\label{pgfplots:speedupVI}
          \addlegendentry{Linear speedup}
          \addlegendentry{$b = 1$}
          \addlegendentry{$b = 3$}
          \addlegendentry{$b = 6$}
    \end{loglogaxis}
    \end{tikzpicture}
    \caption{Scalability of \pk{MatMult} with \pk{MatAIJMKL} and different number of processes.\label{fig:scaling-Ax}}
\end{figure}

@@

decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied a \pk{MatIS}
which stores local unassembled matrices and local to global mappings.
These local matrices are equivalent to the $\{\mathring{A}_i\}_{i=1}^N$

@@

\item whenever the Jacobian is being updated in the \pk{SNESSetJacobian}
function, new unassembled Neumann matrices at the current linearization point are updated as well and supplied to the previously
explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically if the nonliear solver is attached to a

@@
